\section{Results}\label{sec:results}

\subsection{Phase 1: Agreement on the Annotated Pilot Subset}

Phase~1 evaluates agreement on the 30 bug reports for which annotator
consensus labels are available. This is the only phase that supports
agreement claims between the LLM and human annotation.

\subsubsection{Inter-Annotator Agreement}

To verify that the rubric was applied consistently, we computed
Cohen's Kappa between the two human annotators on the pilot subset.
As summarized in Table~\ref{tab:hh_kappa}, agreement was perfect across
all reported dimensions and overall, with raw agreement also equal to
1.000. Within this pilot subset, the rubric was therefore applied
consistently enough to support comparison against the LLM.

\begin{table}[t]
\caption{Human--human agreement on the pilot subset ($N=30$)}
\label{tab:hh_kappa}
\centering
\begin{tabular}{lcc}
\hline
\textbf{Metric} & \textbf{Kappa} & \textbf{Decision} \\
\hline
Overall & 1.0000 & PASS \\
OB      & 1.0000 & PASS \\
EB      & 1.0000 & PASS \\
S2R     & 1.0000 & PASS \\
\hline
Raw agreement & 1.0000 & --- \\
\hline
\end{tabular}
\end{table}

\subsubsection{LLM-vs-Annotator Agreement}

GPT-4o mini was then run on the same 30 pilot reports using
prompt~v1.0 at temperature~0.0. All 30 outputs were parsed
successfully, yielding 0 \texttt{INVALID} rows. Table~\ref{tab:llm_kappa}
reports agreement between the model labels and the annotator consensus.

\begin{table}[t]
\caption{LLM-vs-annotator agreement on the pilot subset ($N=30$)}
\label{tab:llm_kappa}
\centering
\begin{tabular}{lcc}
\hline
\textbf{Metric} & \textbf{Kappa} & \textbf{Decision} \\
\hline
Overall & 0.6183 & PASS \\
OB      & 1.0000 & PASS \\
EB      & 0.6951 & PASS \\
S2R     & 0.4168 & FAIL \\
\hline
Raw agreement & 0.8444 & --- \\
\hline
\end{tabular}
\end{table}

The model reached strong agreement on OB and substantial agreement on
EB, but S2R remained below the accepted threshold of
$\kappa \geq 0.60$. Overall agreement was $\kappa = 0.6183$, which
passed the pilot gate used by the team, although it remained below the
original proposal target of $\kappa \geq 0.70$.

\subsubsection{Output Validity}

The pilot pipeline completed without parse failures. Of the 30 model
responses, all 30 were marked \texttt{VALID} and none were marked
\texttt{INVALID}. This indicates that the prompt and response format
were stable enough to support downstream metric computation.

\subsection{Phase 2: Scale-Up Run on the Full Corpus}

Phase~2 evaluated whether the same pipeline could be executed at the
full-corpus scale of 250 reports. Because annotator consensus labels
for the remaining 220 reports were not available at the time of
writing, this phase supports operational observations only; it does not
support LLM-vs-annotator agreement claims.

The full run used prompt~v3.0-candidate-s2r-relaxed at
temperature~0.0. All 250 responses were parsed successfully, giving an
invalid rate of 0.0\%. The full run therefore demonstrates that the
pipeline remained operationally stable when scaled from 30 to 250
reports.

\begin{table}[t]
\caption{Run validity by phase}
\label{tab:run_validity}
\centering
\begin{tabular}{lccc}
\hline
\textbf{Phase} & \textbf{Reports} & \textbf{VALID} & \textbf{INVALID} \\
\hline
Phase~1 (pilot)   & 30  & 30  & 0 \\
Phase~2 (full)    & 250 & 250 & 0 \\
\hline
\end{tabular}
\end{table}

Taken together, these results show a mixed picture. The model is
operationally reliable as an automated labeling pipeline and shows
useful agreement with annotator consensus overall, but its
dimension-level performance is uneven, with S2R remaining the main
source of disagreement.
