% ==============================================================
% Section 5: Discussion
% Author: Quân (RW)
% File:   paper/sections/05_discussion.tex
% ==============================================================

\section{Discussion}\label{sec:discussion}

\subsection{OB and EB: Where the Gate Performs Well}

The model achieved perfect raw agreement with the annotator consensus
on OB (100\%, Table~\ref{tab:llm_kappa}) and substantial agreement on
EB ($\kappa = 0.695$), both meeting the pre-registered threshold of
$\kappa \geq 0.60$.  As noted in Section~\ref{sec:method:metrics},
OB Kappa is undefined in this sample because every report was labeled
\textit{Sufficient} by both the model and the annotators; we therefore
characterise OB by raw agreement rather than Kappa, and we exclude it
from any averaged value.

EB shows that the model correctly handles the tightened rubric
requirement---that an expectation must be \emph{explicitly stated}
rather than merely guessable from context---introduced in
\texttt{v1.0} (Section~\ref{sec:method:prompts}).  The label
distribution (Table~\ref{tab:label_dist}) shows the model and
annotators assign \textit{Sufficient} at nearly the same rate on EB,
and disagreement is concentrated on the \textit{Ambiguous}/\textit{Missing}
boundary, which is the most subjective call in the rubric.

\subsection{S2R: A Systematic Weakness}

S2R agreement was substantially lower ($\kappa = 0.417$,
Fig.~\ref{fig:kappa_dim}), falling below the threshold in the pilot.
Manual inspection reported in Section~\ref{sec:method:prompts}
identified a systematic cause: \texttt{v1.0} led the model to expect
explicit, numbered reproduction steps, so reports whose reproduction
path was a runnable code snippet or a single command---the dominant
style in \texttt{pandas} and \texttt{scikit-learn}---were labeled
\textit{Missing}, whereas the annotators accepted them as actionable.
The label distribution confirms this directional bias: the model
assigned \textit{Sufficient} to 69 of 90 label slots versus 64 for the
annotators overall, and \textit{Missing} to only 16 versus 23
(Table~\ref{tab:label_dist}), meaning the model under-flagged
incomplete S2R.

Phase~2 used \texttt{v3.0-candidate-s2r-relaxed} to address this
pattern.  However, as stated in Section~\ref{sec:method:prompts}, the
full corpus has no human labels yet, so the effect of \texttt{v3.0} on
agreement is untested.  The Kappa values discussed throughout this
section characterise \texttt{v1.0} only.

\subsection{Annotator Consistency and Rubric Application}

Human--human Kappa was 1.000 across all three dimensions on the
double-labeled subset (Table~\ref{tab:hh_kappa}), indicating that the
two annotators applied the rubric consistently within this sample.
This removes annotation noise as a confounding factor: the
LLM-vs-annotator gap is attributable to the model's interpretation of
the rubric, not to disagreement between the annotators themselves.

It should be noted that perfect agreement between two annotators from
the same team, on a small overlap of 10 issues, does not guarantee the
rubric would be applied identically by a broader or more diverse group.
This limitation is discussed in Section~\ref{sec:threats}.

\subsection{Overall Agreement and Practical Implications}

The overall LLM-vs-annotator Kappa ($\kappa = 0.618$,
Fig.~\ref{fig:kappa_overall}) passed the pilot gate of $\kappa \geq
0.60$ but did not reach the proposal threshold of $\kappa \geq 0.70$.
Phase~2 results confirm that the pipeline is operationally feasible at
scale: 250 reports were screened in approximately 15 minutes at a
total cost of roughly US\$0.086 (US\$0.0003 per report), with 0 of
250 outputs invalid.

In practical terms, the gate is most reliable as a first-pass filter
for OB and EB---flagging reports that clearly lack a stated failure or
expected outcome.  For S2R, the current accuracy is insufficient for
standalone deployment.  A hybrid pipeline in which the model screens
OB and EB while annotators handle S2R on borderline cases better
matches the gate's demonstrated capability.
